---
title: "covid 19"
author: "Diana Reyes"
date: "10/20/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex

$H_0$ The null hypothesis states that dying is independent of one's sex.

$A_0$ The alternative null hypothesis is that dying is not dependent of one's sex.


Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent

```{r}
barchartGC(~result + Sex,data=kRdata,main ="Frequency Bar Chart")
```


```{r}
barchartGC(~result + Sex,data=kRdata,type="percent", main ="Percent Bar Chart")
```
  
### this bar chart is displaying in total frequency terms our data.  Clearly there are a lot less dead people than alive people as is reflected by the sizes of the bars in the display.  

The percentage of females dying is less than males although the percentage of males with the virus is less than females who are infected with the virus. The charts show that the rate of males dying is higher than females by double. 



  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

The results conclude that more females are alive with a 56 percent as oppose to males with a 44 percent rate. Females that are dead is 36 percent as for males it is 64 percent showing a higher death rate for males. 

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The odds of dying if you are a male is 44/1411 and the odds of dying if you are a female is 25/1831

The odd ratio is (44/1411)/(25/1831)

The odds of dying is 2.3 to 1 


## Conclusions

In conclusion we accept the null hypothesis since the results show that more males die as oppose to females. The ratio for males is at 2.3 when for females its at only 1 which prove that males have a higher rate of dying than females.