---
title: "covid 19"
author: "Diana Reyes"
date: "10/20/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent

```{r}
barchartGC(~result + Sex,data=kRdata,main ="Frequency Bar Chart")
```


```{r}
barchartGC(~result + Sex,data=kRdata,type="percent", main ="Percent Bar Chart")
```
  
### this bar chart is displaying in total frequency terms our data.  Clearly there are a lot less dead people than alive people as is reflected by the sizes of the bars in the display.  

The percentage of females dying is less than males although the percentage of males with the virus is less than females. The charts show that the rate of males dying is higher than females.



  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

The results conclude that more females are alive with a 56 percent as oppose to males with a 44 percent. Females that are dead is 36 percent as for males it is 64 percent showing a higher death rate for males. 

## Inferential results 




## Conclusions

